---
title: "ffmpeg issues"
author: "Christof Neumann"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

Load package but don't provide `ffmpeg` location (to be done manually later).

```{r, message = FALSE}
library(avutils)
set_binaries(pathtosox = "~/Documents/utilities/sox", printmessages = FALSE)
```

```{r}
test_binaries()
```

Provide paths to different `ffmpeg` versions (note that the binaries are not included in the repo because the files are pretty large).

```{r}
# set paths to ffmpeg binaries (not on github!)
f2 <- normalizePath("ffmpeg2.8.5/ffmpeg")
f3 <- normalizePath("ffmpeg3.3/ffmpeg")
f4 <- normalizePath("ffmpeg4.2/ffmpeg")
# check versions
system2(command = f2, args = "-version", stdout = TRUE)[1]
system2(command = f3, args = "-version", stdout = TRUE)[1]
system2(command = f4, args = "-version", stdout = TRUE)[1]
```

The videos I use come from [youtube](https://www.youtube.com/watch?v=Yn8j4XRxSck) and were converted using an [online converter](https://www.onlinevideoconverter.com/en/youtube-converter). It is the same video, but downloaded in four different formats (avi, flv, mp4 and wmv). Let's look at their properties:

```{r, echo=FALSE}
# video info
x <- video_info(filein = list.files("testvideo/", full.names = TRUE), pathtoffmpeg = f4)
x$filein <- basename(as.character(x$filein))
kable(x[, -9])
```

So far so good, although here it is already clear that something is going on (i.e. the durations differ between the four videos). Now we have four different videos in different formats. The fact that they all stem from the same source video doesn't matter. What we do next is to extract the audio from each video using the three different `ffmpeg` versions I have available. The audio files go into a folder `tempaudio`.


The command to extract the audio is the following:

`ffmpeg2.8.5/ffmpeg -i 'testvideo/AVI.avi' -y -ar 44100 -ac 1 'tempaudio/ffmpeg2_AVI.wav' -hide_banner`

And here are some info (via `sox`) for the resulting audio, separated by the different video formats. Still, all looking good here, i.e. *within* a video, the audio files seem superficially identical (same number of samples, duration etc...).

```{r, echo=FALSE}
x <- audio_info(list.files("tempaudio", pattern = ".wav", full.names = TRUE))
x$filename <- basename(as.character(x$filename))
x$filesize_sox <- NULL

kable(x[grep("AVI", x$filename), ], row.names = FALSE)
kable(x[grep("FLV", x$filename), ], row.names = FALSE)
kable(x[grep("MP4", x$filename), ], row.names = FALSE)
kable(x[grep("WMV", x$filename), ], row.names = FALSE)
```

\clearpage

Now I run Marvin's yunitator on these 12 audio files. This is run in a virtual machine (following the instructions in the DiViMe documentation[https://divime.readthedocs.io/en/latest/]). The command I used is this:

`vagrant ssh -c 'yunitate.sh data/ english'`

This produces 12 rttm files, corresponding to 12 audio files.

Let's look at the results. Since the recording is MAL-CHI, I drop any FEM that might have been recognized. 

```{r, echo=FALSE}
from <- list.files("tempaudio", pattern = ".rttm$", recursive = TRUE)
to <- file.path("rttmfiles", basename(from))

res <- expand.grid(vid = c("AVI", "FLV", "MP4", "WMV"), ffmpeg = paste0("ffmpeg", 2:4), role = c("CHI", "MAL", "FEM"))
res$dur <- NA
res$n_anno <- NA
# read files by source video
for (i in 1:nrow(res)) {
  x <- to[grepl(pattern = res$vid[i], x = to)]
  x <- x[grepl(pattern = res$ffmpeg[i], x = x)]
  x <- read_rttm(x)
  x <- x[x$tier == as.character(res$role[i]), ]
  res$n_anno[i] <- nrow(x)
  res$dur[i] <- sum(x$duration)
}
```


The first plot shows the cumulative duration of all annotations separated for CHI and MAL.


```{r, echo=FALSE, fig.width=7.2, fig.height=3.3}
par(mfrow = c(1, 2), family = "serif")
pdata <- res[res$role == "CHI", ]
pdata$x <- rep(1:3, each = 4)# + runif(n = 12, -0.05, 0.05)
pdata$sym <- rep(0:3, 3)

plot(0, 0, xlim = c(0.5, 3.5), ylim = c(0, 80), las = 1, axes = FALSE, 
     xlab = "audio extraction from video with", ylab = "cum. duration of annotations", main = "CHI", cex.lab = 0.85)
axis(1, at = 1:3, labels = levels(pdata$ffmpeg), lwd = 0)
axis(2, las = 1)
box()
points(pdata$x, pdata$dur, pch = pdata$sym)
legend("top", ncol = 4, legend = levels(pdata$vid), yjust = 1, xpd = TRUE, pch = 0:3, cex = 0.7)


pdata <- res[res$role == "MAL", ]
pdata$x <- rep(1:3, each = 4)# + runif(n = 12, -0.1, 0.1)
pdata$sym <- rep(0:3, 3)

plot(0, 0, xlim = c(0.5, 3.5), ylim = c(0, 80), las = 1, axes = FALSE, 
     xlab = "audio extraction from video with", ylab = "cum. duration of annotations", main = "MAL", cex.lab = 0.85)
axis(1, at = 1:3, labels = levels(pdata$ffmpeg), lwd = 0)
axis(2, las = 1)
box()
points(pdata$x, pdata$dur, pch = pdata$sym)
legend("top", ncol = 4, legend = levels(pdata$vid), yjust = 1, xpd = TRUE, pch = 0:3, cex = 0.7)

```


The second plot shows the number of annotations detected separated for CHI and MAL.


```{r, echo=FALSE, fig.width=7.2, fig.height=3.3}
par(mfrow = c(1, 2), family = "serif")
pdata <- res[res$role == "CHI", ]
pdata$x <- rep(1:3, each = 4)# + runif(n = 12, -0.05, 0.05)
pdata$sym <- rep(0:3, 3)

plot(0, 0, xlim = c(0.5, 3.5), ylim = c(0, 60), las = 1, axes = FALSE,
     xlab = "audio extraction from video with", ylab = "annotation count", main = "CHI", cex.lab = 0.85)
axis(1, at = 1:3, labels = levels(pdata$ffmpeg), lwd = 0)
axis(2, las = 1)
box()
points(pdata$x, pdata$n_anno, pch = pdata$sym)
legend("top", ncol = 4, legend = levels(pdata$vid), yjust = 1, xpd = TRUE, pch = 0:3, cex = 0.7)


pdata <- res[res$role == "MAL", ]
pdata$x <- rep(1:3, each = 4)# + runif(n = 12, -0.1, 0.1)
pdata$sym <- rep(0:3, 3)

plot(0, 0, xlim = c(0.5, 3.5), ylim = c(0, 40), las = 1, axes = FALSE,
     xlab = "audio extraction from video with", ylab = "annotation count", main = "MAL", cex.lab = 0.85)
axis(1, at = 1:3, labels = levels(pdata$ffmpeg), lwd = 0)
axis(2, las = 1)
box()
points(pdata$x, pdata$n_anno, pch = pdata$sym)
legend("top", ncol = 4, legend = levels(pdata$vid), yjust = 1, xpd = TRUE, pch = 0:3, cex = 0.7)
```

## summary

There are three take-home message from this exercise.

1) When the source for your audio is a video, it matters which version of `ffmpeg` you use to extract the audio. Somewhere between 3.3 and 4.2 there was a change.

2) The consequences of `ffmpeg` version seemingly work differently on different video formats, i.e. in the example here, only audio extracted from the `.avi` produced varying results.

3) And although consistent across different versions of `ffmpeg`, the different video formats produced different results (only `.flv` and `.mp4` produced stable results). In other words, even if you use a single version (the most recent?) of `ffmpeg`, it appears to matter in which file format your video is saved.

To be clear, this is far from a comprehensive study (I only used one DiViMe tool and only one source video), but it suggests that this is likely an issue nevertheless.

Also, I don't know whether `ffprobe` and `soxi` might be better suited to extract information about media files, but this is probbaly a secondary issue if it is one at all.

